---
title: "HABRI_Y1_Exp2_analysis"
author: "Lauren Knose"
date: "2024-09-16"
output: html_document
editor_options: 
  chunk_output_type: console
---
The purpose of this program is to analyze the data generated from the HABRI Year 1 (2021) Experiment 2 (tDOM source experiments). 

# Part 1. Preparing the data

## Step 1. Load the dependent packages and data:

a) loading dependent packages...

```{r}
#install.packages(c("dplyr", "ggplot2", "ggpubr")) #use if first time running code
library(dplyr) #needed for reformatting data
library(ggplot2) #needed for plotting data
library(ggpubr) #needed for adding stats to figures
```

Packages loaded.

b) loading dependent data...

```{r}
data<-read.csv(file="Y1_Exp2_tDOMsource/Y1_Exp2_allData.csv", header=TRUE)
```

Data loaded.

## Step 2. Define and restructure the data, as needed:

a) defining factors...
```{r}
### identify and order labels for source waters ###
unique(data$Lake) 
data$Lake<-factor(data$Lake,levels=c('Maumee','Sandusky', 'Grand')) #source water

### identify and order labels for treatment group ###
unique(data$Treatment) 
data$Treatment<-factor(data$Treatment, levels=c('initial', 
                                                'control', 
                                                'nutrients only', 
                                            'manure', 
                                            'crop',
                                            'wetland', 
                                            'manure + nutrients',
                                            'crop + nutrients',
                                            'wetland + nutrients'),
                       ordered=TRUE)

### identify and order labels for observation time ###
unique(data$Observation) 
data$Observation<-factor(data$Observation, levels=c('pre','post'),
                         ordered=TRUE) 

### identify and order labels for factors ###
unique(data$tDOM_added) 
data$tDOM_added<-factor(data$tDOM_added, levels=c('none','tDOM added'),
                        ordered=TRUE)

unique(data$Nut_added) 
data$Nut_added<-factor(data$Nut_added, levels=c('none','nutrients added'), 
                            ordered=TRUE)

unique(data$tDOM_source)
data$tDOM_source<-factor(data$tDOM_source, levels=c('none', 'wetland','crop', 'manure'),
                         ordered=TRUE)
```

b) restructuring the data...

```{r}
### converting units ###
data<- data %>%
  mutate(NH4_N_ugL=NH4_mgL*1000, NH4_mgL=NULL, #convert mg to ug
         NO3_N_ugL=NO3_mgL*1000, NO3_mgL=NULL,#convert mg to ug
         TNTP_ratio=TN_mgL/TP_mgL) %>% 
  rename(PO4_P_ugL=PO4_ugL)#remove old field

### calculate average initial measurements by source water ###
init<- data %>% #create new data frame 
  filter(Treatment=="initial") %>% #of only initial treatments
  group_by(Lake)  %>% #group by source water
  summarise(pre_Mic_ugL=mean(Microcystin_ugL, na.rm=TRUE), 
            sd_Mic_ugL=sd(Microcystin_ugL, na.rm=TRUE), 
         pre_PC_ugL=mean(PC_ugL, na.rm=TRUE),
         sd_PC_ugL=sd(PC_ugL, na.rm=TRUE),
         pre_Chla_ugL=mean(Chla_ugL, na.rm=TRUE),
         sd_Chla_ugL=sd(Chla_ugL, na.rm=TRUE),
         pre_DOC_mgL=mean(DOC_mgL, na.rm=TRUE),
         sd_DOC_mgL=sd(DOC_mgL, na.rm=TRUE),
         pre_PO4_ugL=mean(PO4_P_ugL, na.rm=TRUE),
         sd_PO4_ugL=sd(PO4_P_ugL, na.rm=TRUE),
         pre_NH4_ugL=mean(NH4_N_ugL, na.rm=TRUE), #mg converted to ug
         sd_NH4_ugL=sd(NH4_N_ugL, na.rm=TRUE), #mg converted to ug
         pre_NO3_ugL=mean(NO3_N_ugL, na.rm=TRUE), #mg converted to ug
         sd_NO3_ugL=sd(NO3_N_ugL, na.rm=TRUE), #mg converted to ug
         pre_TP_mgL=mean(TP_mgL, na.rm=TRUE), 
         sd_TP_mgL=sd(TP_mgL, na.rm=TRUE), 
         pre_TN_mgL=mean(TN_mgL, na.rm=TRUE),
         sd_TN_mgL=sd(TN_mgL, na.rm=TRUE),
         pre_TNTP_ratio=mean(TNTP_ratio, na.rm=TRUE),
         sd_TNTP_ratio=sd(TNTP_ratio, na.rm=TRUE))  
t(init) #prints out the summary by lake of pre measurements

### extract initial values by source water ###
Maumee_preMic<- init$pre_Mic_ugL[init$Lake=="Maumee" & init$pre_Mic_ugL]
Maumee_prePC<- init$pre_PC_ugL[init$Lake=="Maumee" & init$pre_PC_ugL]
Maumee_pre_Chl<- init$pre_Chla_ugL[init$Lake=="Maumee" & init$pre_Chla_ugL]
Sandusky_preMic<- init$pre_Mic_ugL[init$Lake=="Sandusky" & init$pre_Mic_ugL]
Sandusky_prePC<- init$pre_PC_ugL[init$Lake=="Sandusky" & init$pre_PC_ugL]
Sandusky_pre_Chl<- init$pre_Chla_ugL[init$Lake=="Sandusky" & init$pre_Chla_ugL]
Grand_preMic<- init$pre_Mic_ugL[init$Lake=="Grand" & init$pre_Mic_ugL]
Grand_prePC<- init$pre_PC_ugL[init$Lake=="Grand" & init$pre_PC_ugL]
Grand_pre_Chl<- init$pre_Chla_ugL[init$Lake=="Grand" & init$pre_Chla_ugL]

### create new data frame with pre and post measurements ###
data2<- data %>%
  mutate(pre_Mic=ifelse(Lake=="Maumee", Maumee_preMic, 
                        ifelse(Lake=="Sandusky", Sandusky_preMic, Grand_preMic)
                        ),
         pre_PC=ifelse(Lake=="Maumee", Maumee_prePC,
                       ifelse(Lake=="Sandusky", Sandusky_prePC, Grand_prePC)
                       ),
         pre_Chl=ifelse(Lake=="Maumee", Maumee_pre_Chl,
                        ifelse(Lake=="Sandusky", Sandusky_pre_Chl, Grand_pre_Chl)
                        )
         ) %>%
  filter(Treatment!="initial") #remove the initial treatment, since separate field

### verify initials  removed as treatment group and reorder labels ###
unique(data2$Treatment) 
data2$Treatment<-factor(data2$Treatment, levels=c('control', 
                                                'nutrients only', 
                                            'manure', 
                                            'crop',
                                            'wetland', 
                                            'manure + nutrients',
                                            'crop + nutrients',
                                            'wetland + nutrients'),
                       ordered=TRUE)
```

Data restructured.

c) subsetting the data:

```{r}
data_mau<-data2 %>%
  filter(Lake=='Maumee') #creates a new data file with just post measurements in Maumee Bay
data_san<-data2 %>%
  filter(Lake=='Sandusky') #creates a new data file with just post measurements in Sandusky Bay
data_gra<-data2 %>%
  filter(Lake=='Grand') #creates a new data file with just post measurements in Grand Lake
```

Data subset by lake.

## Step 3. Summarize and view the data:

a) checking for NAs...

```{r}
summary(data2$Chla_ugL)  
summary(data2$PC_ugL) 
summary(data2$Microcystin_ugL) 
summary(data2$DOC_mgL) 
summary(data2$TN_mgL) 
summary(data2$TP_mgL) 
summary(data2$NH4_N_ugL) 
summary(data2$NO3_N_ugL) 
summary(data2$PO4_P_ugL) 
summary(data2$TNTP_ratio) 
```

No NAs found.

b) plotting histographs of all response variables...

```{r}
### check the figure size and resolution mins for publication ####
png(file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Figure1.png", width=4,height=4, units="in", res=1200)
SI_fig1<- layout(matrix(c(1,2,3,4,5,6), nrow=2, ncol=3)) 
hist(data2$Microcystin_ugL, 
     main="non-transformed", xlab="Post Microcystin", cex.main=1, cex.lab=1)
hist(log10(data2$Microcystin_ugL), 
     main="log10-transformed", xlab="Post Microcystin", cex.main=1, cex.lab=1)
hist(data2$PC_ugL, 
     main="non-transformed", xlab="Post Phycocyanin", cex.main=1, cex.lab=1)
hist(log10(data2$PC_ugL), 
     main="log10-transformed", xlab="Post Phycocyanin", cex.main=1, cex.lab=1)
hist(data2$Chla_ugL, 
     main="non-transformed", xlab="Post Chlorophyll-a", cex.main=1, cex.lab=1)
hist(log10(data2$Chla_ugL), 
     main="log10-transformed", xlab="Post Chlorophyll-a", cex.main=1, cex.lab=1)
SI_fig1 #prints figure
dev.off() #closes the graphing device
tryCatch({while (T) dev.off()}, error = \(e) NULL) #needed to plot in Plot window again
```

Histographs plotted. All response variables need to be log_10-transformed.

c) summarizing pre and post-measurements across source water...

```{r}
### summarizing the initial conditions by source water ###
summary_init <- data %>%
  filter(Observation=="pre") %>%
  group_by(Lake) %>%
  summarize(avgDOC=mean(DOC_mgL, na.rm=TRUE),
            sdDOC=sd(DOC_mgL, na.rm=TRUE),
            avgTN=mean(TN_mgL, na.rm=TRUE),
            sdTN=sd(TN_mgL, na.rm=TRUE),
            avgTP=mean(TP_mgL, na.rm=TRUE),
            sdTP=sd(TP_mgL, na.rm=TRUE),
            avgTNTP=mean(TNTP_ratio, na.rm=TRUE),
            sdTNTP=sd(TNTP_ratio, na.rm=TRUE),
            avgPO4=mean(PO4_P_ugL, na.rm=TRUE),
            sdPO4=sd(PO4_P_ugL, na.rm=TRUE),
            avgNH4=mean(NH4_N_ugL, na.rm=TRUE),
            sdNH4=sd(NH4_N_ugL, na.rm=TRUE),
            avgNO3=mean(NO3_N_ugL, na.rm=TRUE),
            sdNO3=sd(NO3_N_ugL, na.rm=TRUE),
            avgChla=mean(Chla_ugL, na.rm=TRUE),
            sdChla=sd(Chla_ugL, na.rm = TRUE),
            avgPC=mean(PC_ugL, na.rm=TRUE),
            sdPC=sd(PC_ugL, na.rm=TRUE),
            avgMC=mean(Microcystin_ugL, na.rm=TRUE),
            sdMC=sd(Microcystin_ugL, na.rm=TRUE)) %>%
  ungroup()
write.csv(summary_init, file="Y1_Exp2_tDOMsource/Products/Table1.csv") #prints out the summary by lake of initial microcystin

### summarizing the post measurments across source water ###
png(file="Y1_Exp2_tDOMsource/Products/Figure2.png", 
     width=4,height=3, units="in", res=1200) #pub size and res mins
Fig2<- layout(matrix(c(1,2,3), nrow=1, ncol=3)) 
par(srt=45)
boxplot(data=data2, Chla_ugL~Lake, #create boxplot to show mean Chl-a by Lake
        xlab=NULL, ylab="Chlorophyll-a (ug/L)", las=2)
boxplot(data=data2, PC_ugL~Lake, #create boxplot to show mean Chl-a by Lake
        xlab=NULL, ylab="Phycocyanin (ug/L)", las=2)
boxplot(data=data2, Microcystin_ugL~Lake, #create boxplot to show mean Chl-a by Lake
        xlab=NULL, ylab="Total Microcystins (ug/L)", las=2)
Fig2 #prints figure
dev.off() #closes the graphing device
tryCatch({while (T) dev.off()}, error = \(e) NULL) #needed to plot in Plot window again
```

Post-measurements summarized by source water.

d) summarizing post-measurements across treatments...

```{r}
### summarize and plot all post Chla values ###
summary_chla<- data2 %>% #among the post-measurements
  group_by(Lake, Treatment) %>% #for each lake and treatment
  summarise(avgChla_ugL=mean(Chla_ugL, na.rm=TRUE), #calculate the mean + sd
            sdChla_ugL=sd(Chla_ugL, na.rm = TRUE))
SI_Table2<-print(summary_chla, n=30) #prints the table
write.csv(SI_Table2, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Table2.csv")

SI_fig2<- ggplot(data=data2, 
       mapping=aes(y=Chla_ugL, x=Treatment)) +
  geom_point(aes(shape=tDOM_added, color=Lake)) +
  scale_color_manual(values=c("skyblue3",
                              "olivedrab",
                              "goldenrod")) +
  stat_summary(fun=median, geom="crossbar", width=0.5) +
  facet_wrap(~Lake) +
  labs(x="Treatment Group", y="Chlorophyll-a (ug/L)")+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust=0.5,vjust=0.5),
        legend.position="none")  

SI_fig2
ggsave(SI_fig2, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Figure2_color.png",device="png", width=6, height=6, units="in")

### summarize and plot all post PC values ###
summary_pc<- data2 %>%
  group_by(Lake, Treatment) %>%
  summarise(avgPC_ugL=mean(PC_ugL, na.rm=TRUE),
            sdPC_ugL=sd(PC_ugL, na.rm = TRUE))
SI_Table3<- print(summary_pc, n=30)
write.csv(SI_Table3, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Table3.csv")

SI_fig3<- ggplot(data=data2, 
       mapping=aes(y=PC_ugL, x=Treatment)) +
  geom_point(aes(shape=tDOM_added, color=Lake)) +
  scale_color_manual(values=c("skyblue3",
                              "olivedrab",
                              "goldenrod")) +
  stat_summary(fun.y=median, geom="crossbar", width=0.5)+
  facet_wrap(~Lake) +
  labs(x="Treatment Group", y="Phycocyanin (ug/L)")+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust=0.5,vjust=0.5),
        legend.position = "none") 
SI_fig3
ggsave(SI_fig3, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Figure3_color.png",device="png", width=5, height=5, units="in", dpi=1200)

### summarize and plot all post microcystin values ###
summary_mic<- data2 %>%
  group_by(Lake, Treatment) %>%
  summarise(avgMic_ugL=mean(Microcystin_ugL, na.rm=TRUE),
            sdMic_ugL=sd(Microcystin_ugL, na.rm = TRUE))
SI_Table4<- print(summary_mic, n=30)
write.csv(SI_Table4, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Table4.csv")

SI_fig4<- ggplot(data=data2, 
       mapping=aes(y=Microcystin_ugL, x=Treatment)) +
  geom_point(aes(shape=tDOM_added, color=Lake)) +
  scale_color_manual(values=c("skyblue3",
                              "olivedrab",
                              "goldenrod")) +
  stat_summary(fun.y=median, geom="crossbar", width=0.5)+
  facet_grid(~Lake) +
  labs(x="Treatment Group", y="Total Microcystins (ug/L)")+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust=0.5,vjust=0.5, size=12),
        legend.position = "none") +
  scale_y_continuous(trans='log10') +
  geom_hline(yintercept=0.19, linetype="dashed", color = "red")
SI_fig4
ggsave(SI_fig4, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Figure4_color.png",device="png", width=5, height=5, units="in", dpi=1200)

### post DOC measurements ###
summary_doc<- data2 %>%
  group_by(Lake, Treatment) %>%
  summarise(avgDOC_mgL=mean(DOC_mgL, na.rm=TRUE),
            sdDOC_mgL=sd(DOC_mgL, na.rm = TRUE))
SI_Table5<- print(summary_doc, n =30)
write.csv(SI_Table5, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Table5.csv")

### post TN measurements ###
summary_TN<- data2 %>%
  group_by(Lake, Treatment) %>%
  summarise(avgTN_mgL=mean(TN_mgL, na.rm=TRUE),
            sdTN_mgL=sd(TN_mgL, na.rm = TRUE))
SI_Table6<- print(summary_TN, n = 30)
write.csv(SI_Table6, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Table6.csv")

### post TP measurements ###
summary_TP<- data2 %>%
  group_by(Lake, Treatment) %>%
  summarise(avgTP_mgL=mean(TP_mgL, na.rm=TRUE),
            sdTP_mgL=sd(TP_mgL, na.rm = TRUE))
SI_Table7<- print(summary_TP, n = 30)
write.csv(SI_Table7, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Table7.csv")

### post dissolved measurements by Lake ###
summary<- subset(data2, Lake=="Maumee") %>%
  group_by(Treatment) %>%
  summarise(avg_PO4=mean(PO4_P_ugL, na.rm=TRUE),
            sd_PO4=sd(PO4_P_ugL, na.rm = TRUE),
            avg_NH4=mean(NH4_N_ugL, na.rm=TRUE),
            sd_NH4=sd(NH4_N_ugL, na.rm = TRUE),
            avg_NO3=mean(NO3_N_ugL, na.rm=TRUE),
            sd_NO3=sd(NO3_N_ugL, na.rm = TRUE))
SI_Table8<- print(summary)
write.csv(SI_Table8, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Table8.csv")

summary<- subset(data2, Lake=="Sandusky") %>%
  group_by(Treatment) %>%
  summarise(avg_PO4=mean(PO4_P_ugL, na.rm=TRUE),
            sd_PO4=sd(PO4_P_ugL, na.rm = TRUE),
            avg_NH4=mean(NH4_N_ugL, na.rm=TRUE),
            sd_NH4=sd(NH4_N_ugL, na.rm = TRUE),
            avg_NO3=mean(NO3_N_ugL, na.rm=TRUE),
            sd_NO3=sd(NO3_N_ugL, na.rm = TRUE))
SI_Table9<- print(summary)
write.csv(SI_Table9, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Table9.csv")

summary<- subset(data2, Lake=="Grand") %>%
  group_by(Treatment) %>%
  summarise(avg_PO4=mean(PO4_P_ugL, na.rm=TRUE),
            sd_PO4=sd(PO4_P_ugL, na.rm = TRUE),
            avg_NH4=mean(NH4_N_ugL, na.rm=TRUE),
            sd_NH4=sd(NH4_N_ugL, na.rm = TRUE),
            avg_NO3=mean(NO3_N_ugL, na.rm=TRUE),
            sd_NO3=sd(NO3_N_ugL, na.rm = TRUE))
SI_Table10<- print(summary)
write.csv(SI_Table10, file="Y1_Exp2_tDOMsource/Products/SI_Figures/SI_Table10.csv")
```

Post-measurements summarized by treatments.

# Part 2. Analyze difference in treatments

## Step 1. Comparing controls, tDOM only, and nutrient only groups:

a) running repeated measures ANOVA for Chla...

```{r}
### subset the data ###
SHtest_Chl<- data2 %>% #create new data frame focused on response
  filter(Treatment=="control" | Treatment=="nutrients only"| Treatment =="wetland" | Treatment=="crop"|Treatment=="manure")  #treatments selected

### write the model ###
Chl_anova<- (aov(data=SHtest_Chl, 
               log10(Chla_ugL)~log10(pre_Chl)+Treatment)) 

### check assumptions of the model ###
par(mfrow=c(2,2)) #structure graphs as matrix
SI_Figure5<-plot(Chl_anova) #print model diagnostics
par(mfrow=c(1,1)) #return to default settings

### print the summary statistics from the model ###
summary(Chl_anova) #print the estimated parameters table

### comparison across source waters ###
SI_Fig6a<-ggplot(data=SHtest_Chl, #plots the model
       aes(x=log10(pre_Chl), y=log10(Chla_ugL), color=Treatment)) +
  geom_point(aes(shape=Lake)) + 
  geom_smooth(method="lm", se=TRUE) +
  theme_classic() + #classic publication format
  scale_color_manual(values=c("skyblue3","darkgreen","saddlebrown","gold","darkblue")) +
  labs(x="Pre Chlorophyll-a (ug/L), log10-transformed",
       y="Post Chlorophyll-a (ug/L), log10-transformed",
       color="Treatment", shape="Source Water")
SI_Fig6a

comparisons<- compare_means(data=SHtest_Chl, Chla_ugL ~ Treatment, ref.group = ".all", method="t.test") #name the comparison groups
SI_Fig6b<- ggboxplot(SHtest_Chl, x="Treatment", y="Chla_ugL",
                  add="point", 
                  ylab="Chlorophyll-a (ug/L)") + 
  stat_compare_means(method="anova", label.y=300, label.x=4)+
  stat_compare_means(label="p.signif", method="t.test", ref.group = ".all.") + theme(axis.text.x=element_text(angle=45, hjust=1))
SI_Fig6b
SI_Fig6<- ggarrange(SI_Fig6a,SI_Fig6b, ncol=2, nrow=1)
SI_Fig6
ggsave(SI_Fig6, file="Y1_Exp2_tDOMsource/Products/SI_Figures/Figure6_color.png",device="png", width=6, height=5, units="in", dpi=1200)
```

Chl-a RM-ANCOVA model plotted. 

b) running repeated measures ANOVA for PC...

```{r}
### subset the data ###
SHtest_PC<- data2 %>% #create new data frame focused on response
  filter(Treatment=="control" | Treatment=="nutrients only"| Treatment =="wetland" | Treatment=="crop"|Treatment=="manure")  #treatments selected

### write the model ###
PC_anova<- (aov(data=SHtest_PC, 
               log10(PC_ugL)~log10(pre_PC)+Treatment)) 

### check assumptions of the model ###
par(mfrow=c(2,2))
SI_Figure7<-plot(PC_anova) #print model diagnostics
par(mfrow=c(1,1))

### print the summary statistics from the model ###
summary(PC_anova) #print the estimated parameters table

### comparison across source waters ###
SI_Fig8a<-ggplot(data=SHtest_PC, #plots the model
       aes(x=log10(pre_PC), y=log10(PC_ugL), color=Treatment)) +
  geom_point(aes(shape=Lake)) + 
  geom_smooth(method="lm", se=TRUE) +
  theme_classic() + #classic publication format
  scale_color_manual(values=c("skyblue3","darkgreen","saddlebrown","gold","darkblue")) +
  labs(x="Pre Phycocyanin (ug/L), log10-transformed",
       y="Post Phycocyanin (ug/L), log10-transformed",
       color="Treatment", shape="Source Water")
SI_Fig8a

comparisons<- compare_means(data=SHtest_PC, PC_ugL ~ Treatment, ref.group = ".all.", method="t.test") #name the comparison groups
SI_Fig8b<- ggboxplot(SHtest_Chl, x="Treatment", y="PC_ugL",
                  add="point", 
                  ylab="Phycocyanin (ug/L)") + 
  stat_compare_means(method="anova", label.y=1500, label.x=4)+
  stat_compare_means(label="p.signif", method="t.test", ref.group = ".all.") + theme(axis.text.x=element_text(angle=45, hjust=1))
SI_Fig8b
SI_Fig8<- ggarrange(SI_Fig8a,SI_Fig8b, ncol=2, nrow=1)
SI_Fig8
ggsave(SI_Fig8, file="Y1_Exp2_tDOMsource/Products/SI_Figures/Figure8_color.png",device="png", width=6, height=5, units="in", dpi=1200)
```

PC RM-ANOVA model plotted.

c) running repeated measures ANOVA for microcystin...

```{r}
### subset the data ###
SHtest_MC<- data2 %>% #create new data frame focused on response
  filter(Treatment=="control" | Treatment=="nutrients only"| Treatment =="wetland" | Treatment=="crop"|Treatment=="manure")  #treatments selected

### write the model ###
MC_anova<- (aov(data=SHtest_MC, 
               log10(Microcystin_ugL)~log10(pre_Mic)+Treatment)) 

### check assumptions of the model ###
par(mfrow=c(2,2))
SI_Fig9<-plot(MC_anova) #print model diagnostics
par(mfrow=c(1,1))

### print the summary statistics from the model ###
summary(MC_anova) #print the estimated parameters table

### comparison across source waters ###
Fig10a<-ggplot(data=SHtest_MC, #plots the model
       aes(x=log10(pre_Mic), y=log10(Microcystin_ugL), color=Treatment)) +
  geom_point(aes(shape=Lake)) + 
  geom_smooth(method="lm", se=TRUE) +
  theme_classic() + #classic publication format
  scale_color_manual(values=c("skyblue3","darkgreen","saddlebrown","gold","darkblue")) +
  labs(x="Pre Microcystin (ug/L), log10-transformed",
       y="Post Microcystin (ug/L), log10-transformed",
       color="Treatment", shape="Source Water")
Fig10a

comparisons<- compare_means(data=SHtest_MC, Microcystin_ugL ~ Treatment, ref.group = ".all", method="t.test") #name the comparison groups
Fig10b<- ggboxplot(SHtest_Chl, x="Treatment", y="Microcystin_ugL",
                  add="point", 
                  ylab="Microcystin (ug/L)") + 
  stat_compare_means(method="anova", label.y=60, label.x=4)+
  stat_compare_means(label="p.signif", method="t.test", ref.group = ".all.") + theme(axis.text.x=element_text(angle=45, hjust=1))
Fig10b
Fig10<- ggarrange(Fig10a,Fig10b, ncol=2, nrow=1)
Fig10
ggsave(Fig10, file="Y1_Exp2_tDOMsource/Products/Figure10_color.png",device="png", width=6, height=5, units="in", dpi=1200)
```

Microcystin RM-ANOVA models summarized and plotted.

## Step 2. Compare treatments with 3-way RM-ANCOVA:

a) running three-way, repeated measures ANCOVA of chl-a...

```{r}
### write the model ###
Chl_ancova2<- aov(data=data2, 
               log10(Chla_ugL)~log10(pre_Chl)+tDOM_added*Nut_added*tDOM_source)  

### check assumptions of the model ###
par(mfrow=c(2,2))
SI_Fig11<- plot(Chl_ancova2) #plot diagnostics for model
par(mfrow=c(1,1))

### print the summary statistics from the model ###
summary(Chl_ancova2) #prints the estimated parameters table

### compare within source waters ###
comparisons<- list(c("control", "nutrients only"),
                     c("wetland", "wetland + nutrients"),
                   c("crop", "crop + nutrients"),
                   c("manure", "manure + nutrients"))

Figure3<- ggplot(data=data2, aes(x=tDOM_source, y=Chla_ugL,
                                         color=Nut_added)) +
  geom_point() +
  stat_summary(fun="median", geom="crossbar") +
  theme_classic() +
  scale_color_manual(values=c("darkblue","darkgreen")) +
  theme_classic() +
  facet_wrap(~Lake)+
  theme(axis.text.x=element_text(angle=45, hjust=1, size=12),
        axis.text.y=element_text(size=12),
        strip.text=element_text(size=12),
        axis.title=element_text(size=12),legend.position='top')+
  labs(x="tDOM Source",
       y="Chlorophyll-a (ug/L)",
       color="Nutrients added:") +
  stat_compare_means(label="p.signif", size=8, label.y=500,
                     method="t.test")
Figure3
ggsave(Figure3, file="Y1_Exp2_tDOMsource/Products/Figure3_color.png", device="png", width=6, height=6, units="in", dpi=1200)
```

Chla RM-ANCOVA ran and plotted.

b) running three-way, repeated measures ANCOVA of phycocyanin...

```{r}
### write the model ###
PC_ancova<- aov(data=data2, 
               log10(PC_ugL)~log10(pre_PC)+tDOM_added*Nut_added*tDOM_source)  

### check assumptions of the model ###
par(mfrow=c(2,2))
SI_Fig12<- plot(PC_ancova) #plot diagnostics for model
par(mfrow=c(1,1))

### print the summary statistics from the model ###
summary(PC_ancova) #prints the estimated parameters table

### compare within source waters ###
comparisons<- list(c("control", "nutrients only"),
                     c("wetland", "wetland + nutrients"),
                   c("crop", "crop + nutrients"),
                   c("manure", "manure + nutrients"))

Figure4<- ggplot(data=data2, aes(x=tDOM_source, y=PC_ugL,
                                         color=Nut_added)) +
  geom_point() +
  stat_summary(fun="median", geom="crossbar") +
  theme_classic() +
  scale_color_manual(values=c("darkblue","darkgreen")) +
  theme_classic() +
  facet_wrap(~Lake)+
  theme(axis.text.x=element_text(angle=45, hjust=1, size=12),
        axis.text.y=element_text(size=12),
        strip.text=element_text(size=12),
        axis.title=element_text(size=12),legend.position='top')+
  labs(x="tDOM Source",
       y="Phycocyanin (ug/L)",
       color="Nutrients added:") +
  stat_compare_means(label="p.signif", size=8, label.y=1000,
                     method="t.test")
Figure4
ggsave(Figure4, file="Y1_Exp2_tDOMsource/Products/Figure4_color.png", device="png", width=6, height=6, units="in", dpi=1200)
```

PC RM-ANCOVA ran and plotted.

c) running 3-way RM-ANCOVA of microcystin...

```{r}
### write the model ###
MC_ancova<- aov(data=data2, 
               log10(Microcystin_ugL)~log10(pre_Mic)+tDOM_added*Nut_added*tDOM_source)  

### check assumptions of the model ###
par(mfrow=c(2,2))
SI_Fig13<- plot(MC_ancova) #plot diagnostics for model
par(mfrow=c(1,1))

### print the summary statistics from the model ###
summary(Chl_ancova2) #prints the estimated parameters table

### compare within source waters ###
comparisons<- list(c("control", "nutrients only"),
                     c("wetland", "wetland + nutrients"),
                   c("crop", "crop + nutrients"),
                   c("manure", "manure + nutrients"))

Figure5<- ggplot(data=data2, aes(x=tDOM_source, y=Microcystin_ugL,
                                         color=Nut_added)) +
  geom_point() +
  stat_summary(fun="median", geom="crossbar") +
  theme_classic() +
  scale_color_manual(values=c("darkblue","darkgreen")) +
  theme_classic() +
  facet_wrap(~Lake)+
  theme(axis.text.x=element_text(angle=45, hjust=1, size=12),
        axis.text.y=element_text(size=12),
        strip.text=element_text(size=12),
        axis.title=element_text(size=12),legend.position='top')+
  labs(x="tDOM Source",
       y="Microcystins (ug/L)",
       color="Nutrients added:") +
  stat_compare_means(label="p.signif", size=8, label.y=100,
                     method="t.test")
Figure5
ggsave(Figure5, file="Y1_Exp2_tDOMsource/Products/Figure5_color.png", device="png", width=6, height=6, units="in", dpi=1200)
```

Microcystin RM-ANCOVA model ran and plotted.


d) standardizing microcystin by Chla...

```{r}
### view the data ###
MC_chla<- ggplot(data=data2, aes(x=tDOM_source, y=(Microcystin_ugL/Chla_ugL), color=Nut_added)) +
   geom_boxplot() + 
  facet_wrap(~Lake) +
  theme_classic()+
  theme(axis.text.x=element_text(angle=45, hjust=1, size=12),
        axis.text.y=element_text(size=12),
        strip.text=element_text(size=12),
        axis.title=element_text(size=12), legend.position='none')+
  scale_color_manual(values=c("darkblue","darkgreen")) +
  labs(y="toxin:chl-a")
MC_chla

### calculate means ###
means_MCchla<- data2 %>%
  group_by(Lake) %>%
  summarise(meanMCchla=mean(Microcystin_ugL/Chla_ugL, na.rm=TRUE), 
            sd_MCchla=sd(Microcystin_ugL/Chla_ugL, na.rm=TRUE)) 
means_MCchla

### run ANCOVA ###
MC_chla_ancova<- aov(data=data2, 
               (Microcystin_ugL/Chla_ugL)~(preMic/pre_Chl)+ tDOM_added*Nut_added*tDOM_source) 

summary(MC_chla_ancova)
```

MC:chla analyzed.
